---
title: "Significance testing the class survey - Part I "
subtitle: "Ranjini Grove"
date: "`r Sys.Date()`"
output:    
      html_document:
         toc: no
         toc_float: no
      editor_options:
         chunk_output_type: console
---

Today we will conduct significance tests by simulation to analyze survey responses from STAT 311 students collected at the start of autumn quarter 2024. 

The packages we will be using have been loaded in the code chunk below.

```{r load-packages, include=FALSE}
library(tidyverse)      #umbrella suite
library(tidymodels)     #for infer
library(janitor)
library(skimr)
```

## Data

The class data was collected via a Google form and was downloaded as a CSV file named `classdata_au2024.csv`. The codebook with a description of the variables in the dataset is shown below. The overall data cleaning steps were extremely minimal. I renamed the variables so they were shorter. Additional data cleaning and recoding is done for the variables involved in specific analyses and these will be described in the relevant sections. All the code is included in the Rmarkdown document. 

```{r data-cleaning, include = FALSE}
classdata <- read_csv(file="classdata_au2024.csv")

classdata <- classdata %>%
             rename( class = contains("class standing"),
               credits = contains("credits"),
               diet = contains("vegetarian"),
               sleep = contains("sleep"),
               dating_pref=contains("attractive"),
               sex=contains("sex"),
               study=contains("study"),
               birthday=contains("born"),
               haircut=contains("haircut"),
               shoes=contains("shoes"),
               nervous=contains("nervous"),
               siblings=contains("siblings"),
               which_tire=contains("tire"),
               height=contains("height"),
               speed=contains("fastest"),
               years_in_wa=contains("WA state"),
               travel=contains("visited"),
               salary=contains("earn"),
               breakfast=contains("breakfast"),
               eyewear=contains("eyeglasses"),
               rent=contains("rent"),
               climate_worry = contains("global warming") ) 




```


|Variable |  Description |
| :----:  |  :--------  |
|Timestamp| Self explanatory |
|class    | What is your class standing?|
|credits  | How many credits are you signed up for in Win 2022?|
|vegetarian| Are you a vegetarian?|
|sleep    | How many hours of sleep do you get on average at night?|
|dating_pref| Would you date someone with a great personality even if you did not find them attractive? |
| sex     | What is your biological sex? |
|gender   | To which gender do you most closely identify? |
|study    |How many hours do you study outside of classes (on average) per week? (if this is your first quarter at UW, you may skip this question) |
|birthday | What day of the week were you born? |
|section  | Which quiz section are you in? |
|haircut | What is the most (in dollars) you have paid for a haircut? |
|shoes   | How many pairs of shoes do you own? |
|nervous | How nervous are you about taking STAT 311?|
|siblings | How many siblings do you have? (include step sibs too)|
|which_tire| Suppose you were late to an exam and decided to use the "flat tire" excuse. Which tire would you  pick?|
|height | What is your height (in inches)?|
|years_in_wa | How many years have you lived in WA state? (enter 0 if none, round your answer to a whole number)|
|travel |How many states in the U.S. have you visited?|
|salary| How much money did you earn last year (in dollars )? |
|breakfast| Do you usually eat breakfast in the morning?|
|eyewear | Do you wear eyeglasses or contact lenses?|
|rent | What is your monthly rent?|
|climate_worry| Scientists predict that global warming may have big effects on the polar regions within the next 100 years. One of the possible effects is that the northern ice cap may completely melt. Would this bother you a great deal, some, a little, or not at all if it actually happened?|



## Part 1: which tire would you pick?

The column `which_tire` in `classdata` contains responses to the question "Suppose you were late to an exam and decided to use the flat tire excuse. Which tire would you pick? " The choices given to the student were: left front, right front, left back and right back. Researchers have found that when asked this question, people tend to pick the front tires more often than the back. Does the intro survey data provide evidence to support this claim? Let's do a hypothesis test to answer this question.



### Exercise 1

What proportion of respondents picked a front tire? Let's answer this by first creating a new variable `front_tire` which has two categories: "yes" and "no" depending on whether a front tire was chosen. Save the proportion that picked a front tire in a variable called `obs_prop`.


   
Since the proportion of sample respondents who picked the right front is clearly larger than 50%, it suggests that we can conclude that people tend to pick a front tire more often than the back tires. 

However, the skeptical mindset requires us to consider the possibility that our high percentage could be explained away by sampling variability. This means sample values can and will vary from the population, and we need to understand how much variability we should expected in a sample of our size.

We will therefore need to conduct a significance test to see whether this is in fact a plausible explanation.


### Exercise 2

A significance test pits the following two claims against each other. Which one represents the null and which one is the alternative hypothesis? 

    a. 50% of people will pick a front tire when asked this question.
    
    b. More than 50% of people will pick a front tire when asked this question.

### Exercise 3

Describe in words (no code needed for this part) how you could simulate outcomes under the null hypothesis using a box with two kinds of tickets: one kind says "front" and the other "back". Be sure to mention how many of each kind of ticket are in the box, how many draws you will make, whether they are made with or without replacement and what you will calculate once you have drawn your sample. 


### Exercise 4

Use `tidymodels` to implement your scheme from exercise 3. Fill in the blanks in the code below. 

```{r, eval=F}

set.seed(114)   

null_dist <- classdata %>%
                specify(response = ___, success = "___") %>%
                 hypothesize(null = "___", p = ___ ) %>%
                 generate(reps = 1000, type = "___") %>%
                 calculate(stat="___")



```

### Exercise 4

Write code to visualize the null distribution of the sample proportion. What fraction of the samples gave a value which was at least as large as our observed value? 


Our graph from Exercise 5 shows us we should typically expect to see 50% of sample respondents choosing a front tire if the claim made in the null hypothesis is true. In fact we would only see a value at least as large as our sample value in 21 out of every 1,000 samples. We call 11/1000 the P-value of the test.

So what should we conclude?


Note: we could use the helper functions - visualize, shade_p_value, get_p_value -  from **tidymodels** to perform the visualization and P-value calculation in Exercise 5. Fill in the blanks in the code chunk below and then remove the eval = FALSE option to see how they work.


```{r eval = FALSE}

null_dist %>% visualize() + 
    shade_p_value(obs_stat = ___ , direction = "___")

null_dist %>% get_p_value(obs_stat = ___,
                          direction = "___")
```


**Participation 9: Please input your answers to exercises 1 and 5 in the CANVAS quiz** 

## Part II: Would melting ice caps bother you?

The `climate_worry` column records your response to the question: "Scientists predict that global warming may have big effects on the polar regions within the next 100 years. One of the possible effects is that the northern ice cap may completely melt. Would this bother you a great deal, some, a little, or not at all if it actually happened?"

First let's create and add a new variable - `worried` - which takes the value "yes" if the respondent answered "A great deal" and no if they picked one of the other options. 



### Exercise 1

1. What proportion of respondents said they would be "bothered a great deal"? Write code in the code chunk below to calculate this and save the value in a variable called `obs_prop`.  


### Exercise 2

Based on a large national level survey, it is believed that 66.7% of all Americans would be bothered a great deal by melting ice caps. Our sample value is smaller than this percentage, but not by much. Can we explain the difference as sampling variability? Or does our sample provide strong evidence that this percentage is actually lower than believed?
 Which of the following claims represents the null and which one is the alternative hypothesis? 
 
a. 66.7% of people are bothered a great deal by melting ice caps.
    
b. Less than 66.7% of people are bothered a great deal by melting ice caps.

### Exercise 3

Describe in words (no code needed for this part) how you could simulate outcomes under the null hypothesis using a box with two kinds of tickets: one kind says "bothered a great deal" and the other "not bothered a great deal". Be sure to mention how many of each kind of ticket are in the box, how many draws you will make, whether they are made with or without replacement and what you will calculate once you have drawn your sample. 

### Exercise 4

Use `tidymodels` to implement your scheme from Exercise 4. Use a seed of 252.
Then visualize the null distribution of the sample proportion and shade in the P-value.


### Exercise 5

Calculate the P-value and make a conclusion. 

