---
title: "A grammar of data wrangling: Part 2 "
subtitle: "STAT 311"
author: "YOUR NAME HERE"
date: "`r Sys.Date()`"
output: html_document
---


We now return to the `gapminder` data frame to answer more questions using the data verbs we learned today: count`, group_by` and `summarise`. 
Let's load the packages we will need today. 

```{r setup, include = FALSE}
#load packages and set global options
library(tidyverse)        #umbrella package
library(gapminder)        #dataset is in this package

```

## Warm Up

To learn more about the data set, you can bring up the help file with `?gapminder`. Let's also load the data using `data` and take a look at it in the spreadsheet editor.

Take a look at the variable types  with the `glimpse` function. 

* * * 


## Part 2: count, group_by, summarise

Fill in the blanks in the code provided in each code chunk to accomplish a specific task. Consult the slide deck "Introduction to Data: Data Manipulation" for help. Each code chunk has the `eval` option set to FALSE. Once you have filled in the blanks you may delete this option so your code will run when the document is knitted.

### Exercises


### Exercise 1

In the following code chunk, I have created a new variable called `LifeExp_cat` which categorizes `lifeExp` according to whether it is below 50 or at least 50. Print a frequency distribution of `lifeExp_cat` for the year 2002.


```{r ex1, eval = FALSE}

    gapminder <- gapminder %>% 
                  mutate( lifeExp_cat = ifelse(lifeExp < 50, 
                                               "below 50", 
                                               "at least 50") )

```

### Exercise 2

Summarize the life expectancy `lifeExp` by calculating the mean.

### Exercise 3

Summarize life expectancy `lifeExp` in Africa in 1952 by calculating the mean.

### Exercise 4

Summarize life expectancy `lifeExp` in 1952 for each continent separately by calculating the mean.


### Exercise 5

Summarize life expectancy `lifeExp` in 1952 for each continent separately by calculating the mean and median and sort the results in descending order by median life expectancy. 

### Exercise 6

Create a new variable called `popMil` which is the population in millions. For example, 8425333 is 8.425333 millions.  Then use this new variable to calculate the total population in the world for each year. 


**Participation 2: Please do your calculations for the remaining problems below and then record your answers in CANVAS for the quiz titled Participation 2**



### Exericise 7

Summarise the life expectancy `lifeExp` in 1952 by calculating the mean.

```{r}
gapminder %>% filter(year == 1952) %>% summarise(mean(lifeExp))
```

### Exercise 8

Summarise the life expectancy `lifeExp` in Asia in 1952 by calculating the mean. 
```{r}
gapminder %>% filter(year == 1952, continent == "Asia") %>% summarise(mean(lifeExp))
```

### Exercise 9

Summarise the life expectancy `lifeExp` for each continent in 2002 by calculating the median. Sort your results in descending order.
```{r}
gapminder %>% filter(year == 2002) %>% 
  group_by(continent) %>% summarise(median_life = median(lifeExp)) %>% 
        arrange(desc(median_life))
```

### Exercise 10

Create a new variable called `gdp` which recovers each country's GDP by multiplying the population `pop` and GDP per capita `gdpPercap`. Summarise the `gdp` for each country by calculating the total across all the years. Which 5 countries have the highest totals?

```{r}
gapminder %>%
             mutate(gdp = gdpPercap*pop) %>%
              group_by(country) %>% 
                summarise(total_gdp = sum(gdp)) %>%
                      arrange(desc(total_gdp) )
                 
```
 

